---
title: "Submission 2"
author: "Angelique Cortez"
date: "2024-08-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(1001001)
```

Build a function to create the plots you made for Presentation 1, incorporating any feedback you received on your submission. Your functions should take the following input: (1) the name of the data frame, (2) a list of 1 or more gene names, (3) 1 continuous covariate, and (4) two categorical covariates (10 pts)
Select 2 additional genes (for a total of 3 genes) to look at and implement a loop to generate your figures using the function you created (10 pts)

```{r}
combined_data <- read.csv("~/Dropbox (Dartmouth College)/Dartmouth Classes/Summer2024/Intro to data analysis/data/combined_data_project.csv")
```


#I had issues removing the unknown age with the function, my goal for the next submission is to figure out how by next week
```{r}
# Write a function that can recreate the plots from submission one
generate_plots <- function(data, genes, continuous_covariate, categorical_covariate1, categorical_covariate2) {
  
  for (gene in genes) {
    # Preprocess the data for the gene of interest
    gene_data <- data %>%
      filter(gene == !!gene) %>%
      mutate(
        # Extract age using a regular expression and convert to numeric
        !!continuous_covariate := as.numeric(sub(".*_(\\d{2})y_.*", "\\1", participant_id)),
        # Extract sex using a regular expression (assuming categorical_covariate1 is sex)
        !!categorical_covariate1 := sub(".*_(male|female)_.*", "\\1", participant_id)
      ) %>%
      select(expression, !!sym(continuous_covariate), !!sym(categorical_covariate1), !!sym(categorical_covariate2)) %>%
      distinct()  # Ensure no duplicate rows
    
    # Check if the gene_data has the correct number of data points
    print(paste("Number of data points for gene", gene, ":", nrow(gene_data)))
    
     # Histogram: distribution of expression values
    histogram_plot <- ggplot(gene_data, aes(x = expression)) +
      geom_histogram(binwidth = 10, fill = "blue", color = "black", alpha = 0.7) +
      labs(title = paste("Histogram of", gene, "Gene Expression"),
           x = "Expression Level",
           y = "Frequency") +
      theme_minimal()
    
    print(histogram_plot)
    
    # Scatterplot: expression vs continuous covariate, colored by categorical covariate1
    scatter_plot <- ggplot(gene_data, aes_string(x = continuous_covariate, y = "expression", color = categorical_covariate1)) +
      geom_point(size = 3, alpha = 0.7) +
      scale_x_continuous(breaks = seq(0, 100, by = 10)) +  # Customize x-axis breaks if needed
      labs(title = paste("Scatterplot of", gene, "Expression vs", continuous_covariate),
           x = continuous_covariate,
           y = "Expression Level",
           color = categorical_covariate1) +
      theme_minimal()
    
    print(scatter_plot)
    
    # Boxplot: expression by categorical_covariate1, separated by categorical_covariate2
    box_plot <- ggplot(gene_data, aes_string(x = categorical_covariate1, y = "expression", fill = categorical_covariate2)) +
      geom_boxplot() +
      labs(title = paste("Boxplot of", gene, "Expression by", categorical_covariate1, "and", categorical_covariate2),
           x = categorical_covariate1,
           y = "Expression Level",
           fill = categorical_covariate2) +
      theme_minimal() +
      facet_wrap(as.formula(paste("~", categorical_covariate2))) +
      theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
    
    print(box_plot)
    

  }
}

```

```{r}
# how to use function with three genes
generate_plots(
  data = combined_data, 
  genes = c("ABI1", "ABHD17A", "ABAT"), 
  continuous_covariate = "age", 
  categorical_covariate1 = "sex", 
  categorical_covariate2 = "mechanical_ventilation"
)
```

